### 2013-01-18 article.tex
Ajout article
Former-commit-id: 8a5345302363d84a4be8d28cf92967d10a772a99

\documentclass[a4paper,10pt,twocolumn]{extarticle}
\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[francais]{babel}
\usepackage{textcomp}
\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage{icomma}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[font=sf, labelfont={sf,bf}, margin=1cm]{caption}
\usepackage[french,onelanguage,ruled,lined]{algorithm2e} % pour pseudo code
\usepackage{array}
\usepackage{titlesec, fontspec, titling}

% Specify different font for section headings
\newfontfamily\headingfont[]{GillSans}
\titleformat*{\section}{\LARGE\headingfont}
\titleformat*{\subsection}{\Large\headingfont}
\titleformat*{\subsubsection}{\large\headingfont}
\renewcommand{\maketitlehooka}{\headingfont}

\pretitle{\begin{flushleft}\fontsize{22}{18}}
\posttitle{\par\end{flushleft}\vskip 0.5em}
\preauthor{\begin{flushleft}\Large \lineskip 0.5em}
\postauthor{\par\end{flushleft}}
%\predate{\begin{flushleft}\large}
%\postdate{\par\end{flushleft}}
\setlength{\droptitle}{-1cm}
\title{\textsf{Implémantation d’un système de détection et de reconnaissance faciale}}
\author{Alexandre Bonhomme\\ {\normalsize Université Paul Sabatier, Toulouse}\\
Florent Maufras\\ {\normalsize EFREI, Paris}\\
Timothée Planté\\ {\normalsize EISTI, Pau}\\}% \small{BONA20128906}}
\date{}

\begin{document}
\maketitle
\section{Introduction}

\section{Conclusion}

\begin{thebibliography}{9}
\bibitem{viola01}
  \bsc{Viola P. and Jones M.}, 
  \emph{Rapid object detection using a boosted cascade of simple features}.
  Computer Vision and Pattern Recognition,
  2001.

\end{thebibliography}
\end{document}